Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation

### 2022-04-19 Writing/CEA_3rd_round/replies_rev.rmd
large updates for 3rd round submission

We really appreciate your comments, and we are glad that you expressed your own concerns despite the fact the reviewers were okay with the previous version. This round of revision turned out to be much more important than the previous one. One of your comments was particularity important, and we were able to correct our mistake before this paper is published. Thank you so much for making this article better together with us. We actually made significant changes to section 4.2, which I believe is much improved because it explains clearly why predicting yield levels well at the observed points do not necessarily mean predicting EONR well. Below, you find your comments in black and our replies in blue.

<!-- We really appreciate your comments, and we are glad that you expressed your own concerns despite the fact the reviewers were okay with the previous version. One of your comments was particularity important and we were able to correct our mistake before this paper is published. So, thank you so much. We actually made significant changes to section 4.2, which I believe is much improved because it explains clearly why predicting yield levels well at the observed points do not necessarily mean predicting EONR well. Below, you find your comments in black and our replies in blue. -->

@@

<span style="color:blue">  
Please see Figure D.1 in Appendix D that shows the convergence of CNN for a single simulation round. We cannot show this kind of figure for all the simulation rounds. But, any interested readers can replicate our work as the codes will be accessible publicly. Please also see the figure below. CNN severely underestimates the impact of N and estimated yield response to N is almost flat, but not exactly flat. 
<!-- Please see the figure below showing some examples of yield response function estimated by CNN under "aabbyytt" modeling scenario. Each of the lines indicates yield response curve for different sites.
 -->

@@

<span style="color:blue">  
We are not saying EOIR cannot be estimated by studies mentioned in 28ff. Indeed, any methods can "estimate" EOIR. However, no method can estimate the "true" EONR and we will "never" know true EONRs for real fields. We can only estimate them. This principle is at the core of statistical analysis. We never observe how data is generated (the underlying laws that govern how the world works whether it is physical, biological, or economic), but we can "estimate" (approximate) what they look like. This means that when we try to evaluate how different models perform in predicting EONR using real data, we simply do not have a valid benchmark to compare against. Some may ask, then why don't we use models' yield prediction ability because yield is observable. And this is precisely the dangerous practice that we show can be harmful.

Please note that this is a stark contrast to "yield" (not EOIR) prediction because yields ARE observable and models can be verified. However, that is not the case for EOIR. This is why you need MC simulations where the data generating process (how yield responds to N, and how field characteristics affect yield response to N) is "known" to the researcher. This way, we can test model's ability to predict EONR, not yield. Any journal article that proposes a new statistical method always accompany the theory of why their method works (consistency of their estimation, etc) with MC simulations where the researchers determines how data is generated and apply their method. This is because we can never test the statistical property of a statistical method in principle using the real data. For example, please see @Wager2018a and @oprescu2019orthogonal.

@@

<span style="color:blue">  
Fields studies were used to obtain the plausible range of optimal EONR and the size of error. 
</span>

@@

<span style="color:blue">

We actually realized the results presented in Figure 6 (in the previous round) is completely invalid. As you pointed out the problem with the relationship presented in the figure is basically comparing apples and oranges: comparing across simulation rounds. Figure 8 plots RMSE of yield and EONR predictions by model and simulation round for several simulation rounds. Figure 6 in the previous round was basically mixing RMSE from all the simulation rounds and finding $R^2$ of the two relationship. That is why $R^2$ was extremely low. Therefore, we decided to remove the figure entirely. Also, we no longer claim that yield and EONR predictions are not correlated at all as we realized that it is false. Thank you so much for saving us from claiming the false statement. We conducted simulations with different degrees of error according to your suggestion. **The results are shown in Appendix G**. By looking at Tables 1, 3, G.1, and G.3, you can see that as the error size decreases, both yield and EONR RMSEs decrease. So, in this sense, yield and EONR predictions are indeed correlated as expected.

However, our claim that the models with the highest yield prediction power is not necessarily the best EONR prediciton (or the most profitable) model remains valid. Please note that this claim comes from comparisons **across the models** given the **same** dataset (so the same errors). Since we consider this a very important point for researchers in this field, we now provide much more in-depth discussions about why that is the case under section 4.2. As you suggested, models that can recover the "structure" of the yield response function will perform well in predicting EONR. However, you can predict yield levels at the observed rates (rates that are actually applied in the experiment) without recovering the structure of the yield response function well (predicting yield levels well for wrong reasons). The training process of prediction-oriented ML methods may twist the reasonings of why yield varies in a way that sacrifices its ability to recover yield response function as long as it fits its primary objective of predicting yield levels well at the "observed" nitrogen rate for each site well. Please see our more detailed discussions on this matter in section 4.2.

@@

<span style="color:blue">  
Please see Figure E.1 that plots true yields (without error, epsilon) against predicted yields by model (CF is not presented as it does not predict yield levels as discussed in the article) for a single simulation round. RMSE is presented in the figure as well. As we have 1000 simulations, it is simply infeasible to present figures like this for all the simulation rounds. But, this figure provides a better sense of the value of RMSE and how good the fit is. 
</span>

@@

Please see Figure 2 for illustrations of how yield responds to N site-specifically for several sites within a field. Please note that Figure 6 presents exactly the models' ability to predict the impact of N on yield at different treatments (lowest to the second lowest, second lowest to third lowest, etc). As you can see in the figure, RF underestimates the impact of N on yield severely, which leads to the underestimation of EONR. BRF suffers negative bias as well. On the contrary, CF does not suffer this bias. Of course, the worst of all is CNN, which thinks that the impact of N on yield is almost nascent as you can see in Figure 5. On top of negative bias RF and BRF suffer, their estimation of the impact of N is also less accurate than CF. That is why CF performs better than RF and BRF in profit. Please note that CF does not even estimate Eq. (3). It estimates the change in yield when N is increased from one rate to another rate as discussed in the paper.

<span style="color:blue">

@@

Please see page 3, line 35 for a new list of studies that attempt to estimate EONR. Please note that the previous studies relevant to our study are those that apply ML methods to OFE data with the ultimate intention of estimating site-specific EONR. There is no study that tests the performance of EONR using real data. This is because that is fundamentally impossible as we mentioned above. There are studies that predict field-level (not site-specific within a field) EONR for individual fields one by one using a traditional regression approach, then model EONR using ML methods to see how close their estimated field-level EONRs from the trained ML methods are to the EONR "estimated" for individual fields (e.g., @Qin2018; @Ransom2019 @Wang2021; @wen2021machine). But, please note that the field-level EONRs they compare against are "estimated." Why did they "estimate" EONR for each field? That is because they do not (and will never) know the true EONR for each field. Moreover, the objective of such studies are fundamentally different from ours. Our focus is on site-specific EONR within a field, theirs' are on field-specific EONR. We consider those studies not very relevant. We believe including those studies will simply bloat the introduction.

@@

<span style="color:blue">  
We already discuss our results to the state of the art of the literature. The state of the art is sadly trying to predict yields site-specifically as well as possible. We show that such practice can be misleading and economically harmful when your ultimate goal is to actually help farmers make nitrogen application decisions (predicting EONR well). There is nothing wrong with our approach in yield prediction as evident in Figure E.1<!-- Figure 12 -->. We are using ML methods that have been used in the literature of OFE for site-specific input management. Those studies were published very recently. We added our description about how the size of epsilon affects our results briefly in Appendix G. We added the following at line 25 page 390:

"Finally, the simulation results at a smaller and larger size of the error term are presented in Appendix G as a sensitivity check for interested readers. Basically, qualitative results remain the same."

>>>>

Thank you for your comments. In response to them, we made significant changes that greatly clarified our explanation in section 4.2 of why predicting yield levels well at the observed points need not imply accurate prediction of EONRs. Below, please find your comments in black and our replies in blue.

<!-- We really appreciate your comments, and we are glad that you expressed your own concerns despite the fact the reviewers were okay with the previous version. This round of revision turned out to be much more important than the previous one. One of your comments was particularity important, and we were able to correct our mistake before this paper is published. Thank you so much for making this article better together with us. We actually made significant changes to section 4.2, which I believe is much improved because it explains clearly why predicting yield levels well at the observed points do not necessarily mean predicting EONR well. Below, you find your comments in black and our replies in blue. -->

@@

<span style="color:blue"> 
Please see Figure D.1 in Appendix D, which shows the convergence of CNN for a single simulation round. We cannot show this kind of figure for all the simulation rounds, but any interested readers can replicate our work as the codes will be publicly accessible. Please also see the figure below. CNN severely underestimates the impact of N and estimated yield response to N is almost flat, but not exactly flat. 
</span>

<!-- Please see the figure below showing some examples of yield response function estimated by CNN under "aabbyytt" modeling scenario. Each of the lines indicates yield response curve for different sites.
 -->

@@

<span style="color:blue">  
We are not saying that the studies mentioned in 28ff did not estimate EOIRs. However, "true" EONRs on real fields and sites cannot be known with certainty, but can only be estimated. This principle is at the core of statistical analysis. We never observe the underlying laws that govern how the world works but we can "estimate" those processes. This means that when we try to evaluate how different models perform in predicting EONR using real data, we simply do not have a valid benchmark to compare against. Some may ask, then why don't we use models' yield prediction ability because yield is observable? And this is precisely the dangerous practice that we show can be harmful.

Please note that this is a stark contrast to "yield" (not EOIR) prediction because yields ARE observable and models can be verified. However, that is not the case for EOIR. This is why MC simulations are needed for which the underlying data generating process (how yield responds to N, and how field characteristics affect yield response to N) is "known" to the researcher. This way, we can test a model's ability to predict EONR, not yield. This approach is commonly used in statistical journals when new statistical methods are proposed. This is because we can never test the statistical property of a statistical method in principle using the real data. For example, please see @Wager2018a and @oprescu2019orthogonal.
</span>

@@

<span style="color:blue">  
Some of the authors have significant experience designing, conducting and analyzing hundreds of on-farm precision experiments. The general results of those experiments provided plausible ranges of the optimal EONR and error size. 
</span>

@@

<span style="color:blue">  
We agree that the results presented in Figure 6 of the previous manuscript were invalid and basically compared apples and oranges. The figure was mixing RMSEs from all the simulation rounds and finding the $R^2$ of the two relationships. That is why $R^2$ was extremely low. Therefore, we decided to remove the figure entirely. Also, we no longer make the claim that yield and EONR predictions are completely uncorrelated. Thank you for the heads-up. Figure 8 of the revised manuscript plots RMSE of yield and EONR predictions by model and simulation round for several simulation rounds (simulation round 10, 20, and 30).

In addition, we followed your suggestion to conduct simulations with different degrees of error. **The results are shown in Appendix G**. Examination of Tables 1, 2, G.1, and G.2 reveals that as the error size decreases, both yield and EONR RMSEs decrease. So, in this sense, yield and EONR predictions were indeed correlated, as expected.

We maintain in the revised manuscript our claim that the models with the highest yield prediction power need not provide the best (or the most profitable) EONR prediction. Please note that this claim comes from comparisons **across the models** given the **same** dataset (so the same errors). Since we consider this a very important point for researchers in this field, in the revised manuscript's section 4.2 we provide a more in-depth discussion. As you suggested, models that can recover the "structure" of the yield response function will perform well in predicting EONR. However, yield levels can sometimes be predicted well but for "wrong reasons," making it possible to predict yield levels at the observed rates (rates that are actually applied in the experiment) without recovering the structure of the yield response function. The training process of prediction-oriented ML methods may twist the reasoning behind why yield varies, and do so in a way that sacrifices its ability to recover the yield response function as long as it fits the primary objective of predicting yield levels well at the "observed" nitrogen rate for each site. Please see our more detailed discussions on this matter in section 4.2.
</span>

@@

<span style="color:blue">  
Please see Figure E.1, which plots true yields (without error, epsilon) against predicted yields by model for a single simulation round. (CF is not presented as it does not predict yield levels, as discussed in the article.) RMSE is presented in the figure as well. As we have 1000 simulations, it is infeasible to present figures like this for all the simulation rounds. But, this figure provides a better sense of the value of RMSE and the goodness of fit. 
</span>

@@

<span style="color:blue">  
Please see Figure 2 for illustrations of how yield responds to N site-specifically for several sites within a field. Figure 6 illustrates exactly the models' ability to predict the impact of N on yield at different treatments (lowest to the second lowest, second lowest to third lowest, etc). The figure shows that RF dramatically underestimates the impact of N on yield, which leads to the underestimation of EONR. BRF suffers negative bias as well. CF does not suffer this bias. Of course, CNN is the worst, which thinks that the impact of N on yield is almost nothing<!-- nascent --> as you can see in Figure 6 in the revised manuscript and Figure 1 in this document<!-- Figure 5 -->. In addition to the negative biases from RF and BRF, their estimation of the impact of N is less accurate than CF. That is why CF provides higher profits than do RF and BRF. CF does not even estimate Eq. (3), but rather estimates the change in yield when N is increased discretely, as is discussed in the revised manuscript. 
</span>

@@

<span style="color:blue">
Please see page 3, line 35 for a newly listed study (@barbosa2020risk)  that attempts to estimate EONR. Please note that the previous studies relevant to our study apply ML methods to OFE data with the ultimate intention of estimating site-specific EONR. There is no study that tests the performance of EONR using real data, which is fundamentally impossible, as mentioned above. There are studies that predict field-level (but not site-specific within a field) EONRs for individual fields. Some use a traditional regression approach then model EONR using ML methods to see how close their estimated field-level EONRs from the trained ML methods are to the EONR "estimated" for individual fields (e.g., @Qin2018; @Ransom2019 @Wang2021; @wen2021machine). But, please note that the field-level EONRs they compare against are "estimated." Of course, they had to "estimate" the EONR for each field because they did not (and could not) know the true EONR for each field. Moreover, the objective of such studies are fundamentally different from ours. Our focus is on site-specific EONR within a field, theirs are on field-specific EONR. We consider those studies not very relevant to our purposes, and have left discussion of them out of the introduction for brevity.
</span>

@@

<span style="color:blue">  
We discussed our results with regard to the state of the art literature. (Please see lines 352-357, and lines 442-475.) The state of the art is sadly trying to predict yields site-specifically as well as possible. We show that such practice can be misleading and economically harmful when the ultimate goal is to help farmers make more profitable nitrogen application decisions (predicting EONR well). There is nothing wrong with our approach in yield prediction as evident in Figure E.1<!-- Figure 12 -->. We are using ML methods that have been used in the on-farm experimentation literature for site-specific input management. Those studies were published very recently. We added our description about how the size of epsilon affects our results briefly in Appendix G. We added the following at line 381 page 24:

"*To check for model robustness, simulations were also run assuming smaller and larger error term sizes. Results presented in Appendix G show qualitative results basically unchanged.*"

### 2022-04-19 Writing/appendix.rmd
large updates for 3rd round submission

## Appendix B: Understanding Yield Response Functions Better {-}
Another way to writing Eq. \@ref(eq:eqn3) that may help understanding the yield response Eq. (B.1) better conceptually is as follows:

@@

In this formulation, it is made clear that the base parameters $\alpha$, $\beta$, and $ymax$ are functions of a vector of field characteristics variables $\mathbf{c}=(c_1,\dots,c_K)$, which vary spatially on the field.

Which of $c_1,\dots,c_K$ has an influence vary by the parameters. For example, $\alpha$ represents what the yield would be when no nitrogen is applied in conjunction with $ymax$ because when $N = 0$, $yield = ymax * (1-exp(\alpha))$. So, $\alpha$ is a function of soil organic matter and residual nitrogen among $\mathbf{c}$. $ymax$ determines the maximum yield attainable. So, $ymax$ is a function of other micro nutrients like phosphorus and soil depth among $\mathbf{c}$. $\beta$ reflects how efficiently the crop uses N to create grain mass. So, it may be a function of soil properties that affect the availability of applied N fertilizer to the crop, such as N immobilization, leaching, denitrification, and volatilization among $\mathbf{c}$.

@@

\begin{align}
y_{i,j} 
&=  ymax_{i,j}(1-exp(\alpha_{i,j} + \beta_{i,j} \cdot N_{i}))(1+r^{\varepsilon}_{i,j}) \\
&=  ymax_{i,j}(1-exp(\alpha_{i,j} + \beta_{i,j} \cdot N_{i})) +  ymax_{i,j}(1-exp(\alpha_{i,j} + \beta_{i,j} \cdot N_{i})) \cdot r^{\varepsilon}_{i,j} \\
&= ymax_{i,j}(1-exp(\alpha_{i,j} + \beta_{i,j} \cdot N_{i})) + \varepsilon_{i,j}
\end{align}
, where $\varepsilon_{i,j} = ymax_{i,j}(1-exp(\alpha_{i,j} + \beta_{i,j} \cdot N_{i})) \cdot r^{\varepsilon}_{i,j}$.

@@

```{r cnn-epoch, fig.width=5, fig.height=4, fig.cap = "Figure D.1: An example of CNN performance per training epoch (scenario: aabbyytt)"}
knitr::include_graphics(here("GitControlled/Writing/loss_model4_simID1.png"))
```

@@

Figure E.1 plots the predicted yields against the true yield without error in one of the one thousand simulation rounds by ML method and modeling scenario. The predicted points clustered around the red 1-to-1 line, meaning that the prediction-oriented ML methods predicted yields well.

However, as mentioned in section 4.2, please note that even though the yield prediction accuracy evaluated at the actually applied N rate is , it does not necessary mean good performance in predicting yield “response” function which requires counterfactual prediction.

@@

Table D.1 shows the mean RMSE of the EONR estimation and $\hat{\pi}_{def}$ over the one thousand simulations using the training datasets as testing datasets. RF and BRF performs slightly better compared to the case when they are evaluated against the separate testing datasets. However, the relative ranking of the models is the same.

@@

Table D.2 presents the training accuracy of yield prediction: namely, the yield prediction accuracy when the trained model is evaluated using the training datasets as opposed to separate testing datasets.

@@

Using those new yield datasets, we conducted one hundred rounds of simulations respectively. The results of EONR estimation and yield prediction are shown in Table G.1 and G.2. Table G.3 summarizes the relationship between yield and EONR prediction RMSE.

>>>>

## Appendix B: Understanding Yield Response Functions Better {-}
Another way to write Eq. \@ref(eq:eqn3) that may help to better understand the yield response of Eq. (B.1) is as follows:

@@

In this formulation, it is made clear that the base parameters $\alpha$, $\beta$, and $ymax$ are functions of a vector of field characteristics variables $\mathbf{c}=(c_1,\dots,c_K)$, which vary spatially on the field,

$c_1,\dots,c_K$ has an influence that varies by parameter. For example, $\alpha$ represents what the yield would be when no nitrogen is applied in conjunction with $ymax$ because when $N = 0$, $yield = ymax * (1-exp(\alpha))$. So, $\alpha$ is a function of soil organic matter and residual nitrogen among $\mathbf{c}$. $ymax$ determines the maximum yield attainable. So, $ymax$ is a function of other micro nutrients like phosphorus and soil depth among $\mathbf{c}$. $\beta$ reflects how efficiently the crop uses N to create grain mass. It may be a function of soil properties that affect the availability of applied N fertilizer to the crop, such as N immobilization, leaching, denitrification, and volatilization among $\mathbf{c}$.

@@

\begin{align}
y_{i,j} 
&=  ymax_{i,j}(1-exp(\alpha_{i,j} + \beta_{i,j} \cdot N_{i}))(1+r^{\varepsilon}_{i,j}) \\
&=  ymax_{i,j}(1-exp(\alpha_{i,j} + \beta_{i,j} \cdot N_{i})) +  ymax_{i,j}(1-exp(\alpha_{i,j} + \beta_{i,j} \cdot N_{i})) \cdot r^{\varepsilon}_{i,j} \qquad \text{(C.1)}\\
&= ymax_{i,j}(1-exp(\alpha_{i,j} + \beta_{i,j} \cdot N_{i})) + \varepsilon_{i,j}
\end{align}
, where $\varepsilon_{i,j} = ymax_{i,j}(1-exp(\alpha_{i,j} + \beta_{i,j} \cdot N_{i})) \cdot r^{\varepsilon}_{i,j}$.

@@

```{r cnn-epoch, fig.width=5, fig.height=4, fig.cap = "Figure D.1: An example of CNN performance per training epoch (scenario: aabbyytt)"}
knitr::include_graphics(here("GitControlled/Writing/Loss_epoch.eps"))
```

@@

Figure E.1 plots the predicted yields against the true yield in one of the one thousand simulation rounds by ML method and modeling scenario. The RMSEs are presented in the figure as well. Since CF-base does not predict yield, the figure only reports the results from RF, BRF and CNN. The predicted points clustered around the red 1-to-1 line, meaning that the prediction-oriented ML methods predicted yield levels well. <!-- However, as mentioned in section 4.2, even though the yield prediction accuracy evaluated at the actually applied N rate is good, it does not necessarily mean good performance in predicting yield “response” function which requires counterfactual prediction.   -->

@@

Table F.1 shows the mean RMSE of the EONR estimation and $\hat{\pi}_{def}$ over the one thousand simulations using the training datasets as testing datasets. RF and BRF performs slightly better compared to the case when they are evaluated against the separate testing datasets. However, the relative ranking of the models is the same.

@@

Table F.2 presents the training accuracy of yield prediction: namely, the yield prediction accuracy when the trained model is evaluated using the training datasets as opposed to separate testing datasets.

@@

Using those new yield datasets, we conducted one hundred rounds of simulations respectively. The results of EONR estimation and yield prediction are shown in Table G.1 and G.2. Table G.3, which summarize the relationship between yield and EONR prediction RMSE.
